Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-04-26 doc/paper.Rmd
EERs and AUCs

@@

\hh{
We started out trying to assess the suitability of the Chumbley Score's suitability for matching striae on bullet lands. Besides using recommended defaults, we also proposed ways to optimize parameters: based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile, we found  for bullets the smaller coarseness value of $c = 0.15$ to be suitable. 
Optimized window sizes were found based on cross-validation to minimize type 2 error rates. 
Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of XXX for an optimized window size of YYY - which is considerably higher than the error rates achieved on matching toolmarks. 
Unfortunately, method CS1 also has a high rate of failed tests -- situations, in which the algorithm does not provide a result, due to the way different-shift pairs are constructed. 
XXX
The algorithm CS2 introduced here is constructed in a way that reduces the number of failures by a magnitude of XXX. While reducing the failure rate, the algorithm also shows an increase in the power of the test. Type II error rates reach a minimum of XXX for an optimized window size of YYY. While significantly reduced over CS1, CS2 still has type 2 error rates on bullet lands that are  higher than the error rates achieved on the -much larger- toolmarks.
However, bullets usually have multiple lands - in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet. We might be able to get more power out of the test by adapting CS2 to deal work in a bullet-to-bullet comparison.
}

\hh{
Fix in CS2 will also improve power for matching toolmarks. 
}

>>>>

```{r eer, fig.height = 3.25, fig.width = 8, out.width='\\textwidth', fig.cap=""}
auc <- function(tpr, fpr){
  # from http://blog.revolutionanalytics.com/2016/11/calculating-auc.html
  # inputs already sorted, best scores first 
  dfpr <- c(diff(fpr), 0)
  dtpr <- c(diff(tpr), 0)
  sum(tpr * dfpr) + sum(dtpr * dfpr)/2
}

eer <- function(tpr, fpr) {
  diff = abs(fpr-1+tpr)
  idx <- which.min(diff)
  EER <- 0.5*(fpr[idx]+1-tpr[idx])
  EER
}

EER <- function(score, match) {
  bstats <-data.frame(score, match)
  bstats <- bstats %>% arrange(desc(score))

  bstats <- bstats %>% mutate(
    tpr=cumsum(match)/sum(match), 
    fpr=cumsum(!match)/sum(!match)
  ) %>% unique()

  with(bstats, eer(tpr, fpr))
}

AUC <- function(score, match) {
  bstats <-data.frame(score, match)
  bstats <- bstats %>% arrange(desc(score))

  bstats <- bstats %>% mutate(
    tpr=cumsum(match)/sum(match), 
    fpr=cumsum(!match)/sum(!match)
  ) %>% unique()

  with(bstats, auc(tpr, fpr))
}

roc.summary <- cvcs %>% group_by(wo, wv, method) %>%
  summarize(
    eer = EER(1-p_value, match),
    auc = AUC(1-p_value, match)
  )
p1 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
  ggplot(aes( x = wo, y=eer, shape=factor(wv))) +
  geom_point() +
  facet_wrap(~method, labeller="label_both") + 
  theme_bw() +
  scale_shape_discrete("Size of \nvalidation window") +
  ylab("Equal Error rate (EER)") +
  xlab("Size of optimization window in pixels") +
  theme(legend.position = "none")

p2 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
  ggplot(aes( x = wo, y=auc, shape=factor(wv))) +
  geom_point() +
  facet_wrap(~method, labeller="label_both") + 
  theme_bw() +
  scale_shape_discrete("Size of \nvalidation window") +
  ylab("Area under the curve (AUC)") +
  xlab("Size of optimization window in pixels")+
  theme(legend.position = "none")
grid.arrange(p1, p2)
```

@@

\hh{
We started out trying to assess the suitability of the Chumbley Score's suitability for matching striae on bullet lands. Besides using recommended defaults, we also proposed ways to optimize parameters: based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile, we found  for bullets the smaller coarseness value of $c = 0.15$ to be suitable. 
Optimized window sizes were found based on cross-validation to minimize type 2 error rates. 
Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of XXX for an optimized window size of YYY - which is considerably higher than the error rates achieved on matching toolmarks. 
Unfortunately, method CS1 also has a high rate of failed tests -- situations, in which the algorithm does not provide a result, due to the way different-shift pairs are constructed. 
XXX
The algorithm CS2 introduced here is constructed in a way that reduces the number of failures by a magnitude of about XXX. While reducing the failure rate, the algorithm also shows an increase in the power of the test. Type II error rates reach a minimum of XXX for an optimized window size of YYY. While significantly reduced over CS1, CS2 still has type 2 error rates on bullet lands that are  higher than the error rates achieved on the -much larger- toolmarks.
However, bullets usually have multiple lands - in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet. We might be able to get more power out of the test by adapting CS2 to deal work in a bullet-to-bullet comparison.
}

\hh{
Fix in CS2 should also improve power for matching toolmarks. 
}
